### 2025-04-14 LICS25/rebuttal-600words.md
move past to LICS 2025 folder

# Overall Response

We express our gratitude for the valuable feedback received from reviewers. 
We will carefully address the points raised and implement the necessary changes to our paper.

First, we address concerns surrounding lazy normalization. 
The classical algorithm guarantees a full traversal of the coalgebra during liveness checking prior to bisimulation checking, our algorithm invokes bisimulation and only traverses the reachable states of the discrepancies.
Therefore, lazy liveness checking consistently visits fewer states, even for positive results. 
Additionally, lazy liveness checking enables on-the-fly coalgebra generation using derivatives.

While our algorithms seem intuitive, our proofs are not trivial. 
Specifically, the simple proofs are enabled by the design of greedy bisimulation, which took several iterations.
Without this middle ground, verifying Theorem 22 would be challenging, especially given that neither the input coalgebras nor their bisimulation are assumed to be finite.

Finally, we emphasize that our contribution extends beyond bisimulation of coalgebras.
We also presented two methods for generating coalgebras from expressions; and proved their correctness, finiteness, and correspondence.

Regarding our experiments, we have uploaded new tests to our repository, and will provide the code and documentation for test generation.
The performance preserves the efficiency of previous versions, as neither trivial nor dead tests provide performance benefit in our implementation. 
Additionally, we already have several conceptual demonstrations of performance improvements in the paper.

# Reviewer Questions

### Reviewer 1

> Where are the benchmarks from?

It is generated by https://github.com/c-cube/qcheck, based on congruence, symmetry, and some algebraic rules of GKAT (excluding rules that will be "trivial" tests, as named by Reviewer 3).

### Reviewer 2

Our approach diverges from previous works on symbolic transducers in several aspects:
- Our on-the-fly method is tailored for normalization, a challenge that is not shared by symbolic transducers. 
- The signature of GKAT differs from traditional symbolic transducers. Unlike typical accepting or rejecting states, a GKAT state can accept or reject based on different input atoms. This results in a semantics where a trace ends in an atom, representing the output program state.

While symbolic bisimulation is a valuable contribution, it is not our sole contribution. 
We spent most of the paper developing an on-the-fly bisimulation algorithm and two distinct method to construct symbolic GKAT coalgebra from expressions, and proved their desirable properties.

# Detailed Response

### Reviewer 1

> it's unclear to me what section III is for

The correctness result of section IV is based on that of section III.

### Reviewer 2

> can you say a bit more about SymKAT

SymKAT is a symbolic equivalence checker for KAT; on the other hand, our algorithm focuses on the GKAT fragment.
This specialization allows us to produce smaller coalgebra and utilize efficient SAT solver.
Our comparison demonstrated that GKAT equivalence can be checked more efficiently using our specialized solver.

### Reviewer 3

> Theorem 23 is not the proper tool to justify the correctness of the algorithm

Union-find structures represent equivalence relations. 
By employing union-find instead of a set of pairs, our algorithm searches for a greedy bisimulation that is also an equivalence relation. 
While the correctness of greedy bisimulation is established (Theorem 21), Theorem 23 justifies finding a greedy bisimulation that is also an equivalence relation is correct.

> p7c2 [...] the liveness analysis would be faster

Consider two bisimilar states in a GKAT coalgebra with no dead state. 
The classical algorithm would perform a liveness check, iterating through all the states; then proceed with bisimulation, iterating through all the states again. 
Our algorithm simply performs the bisimulation, bypassing the liveness check altogether.